---
title: "Tutorial_3"
output: html_document
date: "2024-06-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install the ggplot2 package
or better, the tidyverse meta-package
install a fork of the medicaldata package from GitHub

```{r medicaldata}
devtools::install_github("cbdrh-hdat9800/medicaldata")
```

## load the libraries
## Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r librarys, echo=FALSE}
library(tidyverse)
library(medicaldata)
```
Create a ggplot
```{r plot, echo=FALSE}
ggplot(data = diabetes, aes(x = age, y = bmi)) + 
  geom_point()

```


```{r plot2, echo=FALSE}
ggplot(data = diabetes, aes(x = age, y = bmi, color = diabetes_5y)) + 
  geom_point(size = 1, alpha = 0.7) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "darkred") +
  labs(title = "BMI vs Age in Diabetes Dataset",
       subtitle = "Colored by Diabetes Status (5 Years)",
       x = "Age",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )
```

Experiment 1
make a scatter plot of bmi versus glucose_mg-dl
hint: it may not work as you expect, but we’ll fix that!

```{r Scatter Plot, echo=FALSE}
ggplot(data = diabetes, aes(x = `glucose_mg-dl`, y = bmi)) + 
  geom_point() +
  labs(title = "Scatter Plot of BMI vs Glucose Levels",
       x = "Glucose (mg/dl)",
       y = "BMI")
```
Enhanced Scatter Plot
```{r Scatter Plot2, echo=FALSE}
ggplot(data = diabetes %>% drop_na(bmi, `glucose_mg-dl`), aes(x = `glucose_mg-dl`, y = bmi, color = diabetes_5y)) + 
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Scatter Plot of BMI vs Glucose Levels",
       subtitle = "Colored by Diabetes Status (5 Years)",
       x = "Glucose (mg/dl)",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )
```
We need to rename (or quote) that column with - in it!

```{r Scatter Plot2, echo=FALSE}
# Rename the column to avoid issues
diab <- diabetes %>%
  rename(glucose_mg_dl = `glucose_mg-dl`)


ggplot(data = diab %>% drop_na(bmi, glucose_mg_dl), aes(x = glucose_mg_dl, y = bmi, color = diabetes_5y)) + 
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Scatter Plot of BMI vs Glucose Levels",
       subtitle = "Colored by Diabetes Status (5 Years)",
       x = "Glucose (mg/dl)",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )
```
make a scatter plot of bmi versus diabetes_5y

```{r Scatter Plot of bmi versus diabetes_5y, echo=FALSE}
ggplot(data = diab %>% drop_na(bmi, diabetes_5y), aes(x = diabetes_5y, y = bmi, color = diabetes_5y)) + 
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(title = "Scatter Plot of BMI vs Diabetes Status",
       x = "Diabetes Status (5 Years)",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )
```
```{r aesthetic mapping, echo=FALSE}
ggplot(data = diab) + 
  geom_point(mapping = aes(x = age, y = bmi, colour = diabetes_5y))
```



```{r Enhanced Scatter Plot with Average BMI Lines, echo=FALSE}
# Calculate Mean BMI for Each Group
mean_bmi_pos <- diabetes %>% filter(diabetes_5y == "pos") %>% summarize(mean_bmi = mean(bmi, na.rm = TRUE)) %>% pull(mean_bmi)
mean_bmi_neg <- diabetes %>% filter(diabetes_5y == "neg") %>% summarize(mean_bmi = mean(bmi, na.rm = TRUE)) %>% pull(mean_bmi)


ggplot(data = diabetes, aes(x = age, y = bmi, color = diabetes_5y)) + 
  geom_point(size = 1, alpha = 0.7) + 
  geom_hline(yintercept = mean_bmi_pos, linetype = "dotted", color = "#00BFC4", size = 1) +
  geom_hline(yintercept = mean_bmi_neg, linetype = "dotted", color = "#F8766D", size = 1) +
  labs(title = "BMI vs Age in Diabetes Dataset",
       subtitle = "Colored by Diabetes Status (5 Years)",
       x = "Age",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )

```

Add a one-dimensional facetting
write some code to add faceting by the diabetes_5y variable to a scatter plot of age for the x-axis and bmi for the y-axis

hint: look at facet_wrap() in ggplot2

use ?facet_wrap at the R Console prompt of visit the ggplot2 documentation web site
```{r Scatter Plot with Faceting, echo=FALSE}
ggplot(data = diabetes, aes(x = age, y = bmi, color = diabetes_5y)) + 
  geom_point(size = 1, alpha = 0.7) + 
  geom_hline(yintercept = mean_bmi_pos, linetype = "dotted", color = "#00BFC4", size = 1) +  # Using color for "pos" points
  geom_hline(yintercept = mean_bmi_neg, linetype = "dotted", color = "#F8766D", size = 1) +  # Using color for "neg" points
  facet_wrap(~ diabetes_5y) +  # Faceting by diabetes_5y
  labs(title = "BMI vs Age in Diabetes Dataset",
       subtitle = "Faceted by Diabetes Status (5 Years)",
       x = "Age",
       y = "BMI",
       color = "Diabetes Status",
       caption = "Source: Diabetes Dataset") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "bottom"
  )


```

































