---
title: "Tutorial_5"
author: "Zhenyu_Zhang(z5037788)"
date: "2024-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data set}
# install.packages("gganimate")
# install.packages("gifski")
# install.packages("av")
# install.packages("plotly")
# install.packages("gifski")

library(tidyverse)
library(medicaldata)
library(ggplot2)
library(gganimate)
library(plotly)
library(dplyr)
library(gifski)

# Load and prepare the data
data("esoph_ca", package = "medicaldata")
# Check the structure and column names of the dataset
glimpse(esoph_ca)
```

```{r relative risk}
# Prepare the data
esoph_ca <- esoph_ca %>%
  mutate(total = ncases + ncontrols,
         relative_risk = ncases / total,
         agegp_numeric = as.numeric(as.factor(agegp)))  # Convert age groups to a numeric format

```

Using Plotly
plotly is great for creating interactive web-based graphs. Here’s how you can create an interactive plot with plotly using your esoph_ca dataset. We'll make a plot where users can choose different age groups to see changes in relative risk across alcohol consumption categories.
```{r plot set}
# Create an interactive plot
p <- esoph_ca %>%
  plot_ly(x = ~alcgp, y = ~relative_risk, color = ~agegp, colors = RColorBrewer::brewer.pal(8, "Dark2"),
          type = 'bar', text = ~paste("Relative Risk: ", relative_risk)) %>%
  layout(barmode = 'group',
         title = "Relative Risk by Alcohol Consumption and Age Group",
         xaxis = list(title = "Alcohol Consumption Group"),
         yaxis = list(title = "Relative Risk"),
         legend = list(title = list(text = "Age Group")),
         hovermode = "closest")

# Render the plot
p

```

shiny is another powerful option for building interactive web applications directly in R. Here's a simple example of a Shiny app that allows users to select an age group and dynamically updates the plot based on the selection:





```{r plot set2}
library(shiny)
library(ggplot2)
library(dplyr)
library(medicaldata)

# Load data
data("esoph_ca", package = "medicaldata")

# Define the UI
ui <- fluidPage(
  titlePanel("Interactive Visualization of Esophageal Cancer Risk"),
  sidebarLayout(
    sidebarPanel(
      # Dropdown menu for alcohol consumption levels
      selectInput("alcLevel", "Select Alcohol Consumption Level:",
                  choices = c("0-39g/day", "40-79g/day", "80-119g/day", "120+g/day"))
    ),
    mainPanel(
      plotOutput("riskPlot")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  output$riskPlot <- renderPlot({
    # Filter data based on the dropdown input
    filtered_data <- esoph_ca %>%
      filter(alcgp == input$alcLevel) %>%
      group_by(agegp) %>%
      summarise(relative_risk = sum(ncases) / sum(ncases + ncontrols))

    # Plot
    ggplot(filtered_data, aes(x = agegp, y = relative_risk, fill = agegp)) +
      geom_col() +
      labs(title = "Relative Risk by Age Group",
           subtitle = paste("Alcohol consumption:", input$alcLevel),
           x = "Age Group", y = "Relative Risk") +
      theme_minimal()
  })
}

# Run the application
shinyApp(ui = ui, server = server)

```



```{r plot set3}
library(ggplot2)
library(gganimate)

ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  # Here comes the gganimate code
  transition_states(
    gear,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')
```



```{r plot set4}
library(gapminder)

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```



```{r plot set5}

```



```{r plot set6}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species, group = seq_along(Species))) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)
```



```{r plot set7}

```

